Improved Approximation Algorithms for the Maximum Carpool Matching Problme

### 2017-02-10 src/abstract.tex
merge dror

The \carpool problem is a star packing problem in directed graphs.
Formally, given a directed graph $G = (V, A)$, a capacity function $c:
V \rightarrow \N$, and a weight function $w : A \rightarrow \R $, a
feasible \emph{carpool matching} is a subset of arcs, $M \subseteq A$,
such that every $v \in V$ satisfies:
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A vertex $v$ for which $\dout[M](v) = 1$ is a \emph{passenger}, and a
vertex for which $\dout[M](v) = 0$ is a driver who has $\din[M](v)$
passengers.  In the \carpool problem the goal is to find a matching
$M$ of maximum total weight.

The problem arises when designing an online carpool service, such as
Zimride~\cite{zimride}, which tries to connect between passengers and
drivers based on (arbitrary) similarity function.  The problem is
known to be NP-hard, even for uniform weights and infinite capacities.
%without capacity constraints.

The \gcp, is an extension of the \carpool where each vertex represents a
(unsplittable) group of passengers.
Each group might have a different size.  
Formally, each vertex $u \in V$ has a size $s(u) \in \N$,
and the constraint $\din(v) \leq c(v)$ is replaced with
$\sum_{u:(u,v) \in M} s(u) \leq c(v)$.

We show that the \carpool can be formulated as an unconstrained
submodular maximization problem, thus it admits a
$\frac{1}{2}$-approximation algorithm.
We show that the same formulation does not hold for the \gcp,
nevertheless, we present a $\frac{1}{2} - \varepsilon$-approximation algorithm
for the \gcp as well.
For the unweighted variant of
the problems when the maximum possible capacity, $C$, is fixed, we give
a $(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm.
We also show that the problem remains APX-hard even if $C \leq 3$.

>>>>

The \carpool problem is a star packing problem in directed graphs.
Formally, given a directed graph $G = (V, A)$, a capacity function $c:
V \rightarrow \N$, and a weight function $w: A \rightarrow \R^+$, a
feasible \emph{carpool matching} is a subset of arcs, $M \subseteq A$,
such that every $v \in V$ satisfies:%
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A vertex $v$ for which $\dout[M](v) = 1$ is a \emph{passenger}, and a
vertex for which $\dout[M](v) = 0$ is a \emph{driver} who has
$\din[M](v)$ passengers.  In the \carpool problem the goal is to find
a matching $M$ of maximum total weight.
%
The problem arises when designing an online carpool service, such as
Zimride~\cite{zimride}, which tries to connect between passengers and
drivers based on a similarity function.  The problem is known to be
NP-hard, even for uniform weights and infinite capacities.
%
The \gcp problem, is an extension of the \carpool where each vertex
represents an unsplittable group of passengers.  Formally, each vertex
$u \in V$ has a size $s(u) \in \N$, and the constraint $\din(v) \leq
c(v)$ is replaced with $\sum_{u:(u,v) \in M} s(u) \leq c(v)$.

We show that \carpool can be formulated as an unconstrained submodular
maximization problem, thus it admits a $\frac{1}{2}$-approximation
algorithm.  We show that the same formulation does not work for \gcp,
nevertheless, we present a $(\frac{1}{2} - \eps)$-approximation
algorithm for \gcp.
%
For the unweighted variant of both problems when the maximum possible
capacity, $\cmax \eqdf \max_{v \in V} c(v)$, is bounded by a constant,
we provide a $(\half + \inv{2\cmax} - \eps)$-approximation algorithm.
We also show that the problem remains APX-hard, even if $\cmax \leq
3$.

### 2017-02-10 src/group-apx.tex
merge dror

We now argue that this algorithm can be generalized to solve the group carpool
matching problem.
The main concern when trying to adopt the algorithm to the group carpool
matching is how to determine if a vertex can be improved.
Observe that Remark~\ref{rem:improve} does not hold anymore.
If the maximum possible capacity is bounded, one can test for improvement of
$v$ by considering all the subsets of edges intersecting $v$. 
When the capacity is unbounded, however, it is easy to see that efficient
way to test improvement implies a solution to the Knapsack problem.
Indeed, if this is the case, one can test for improvement by using a known FPTAS
for Knapsack, this is in the cost of additional $\varepsilon$ reduction in the
approximation ratio of the algorithm.

>>>>

We now argue that the local improvement algorithm from
Section~\ref{sec:improve} can be generalized to solve \gcp.
%
The main concern when trying to adopt the algorithm to \gcp is how to
determine if a vertex can be improved.
%
With \carpool, if weights are polynomially-bounded, it was enough to
consider in the incoming arcs to a vertex $v$ in a non-increasing
order of $\delta_M$ (See proof of Theorem~\ref{thm:improve-bounded}).
This does not work anymore, since \gcp we have sizes.  In fact, given
$v$, finding the best star with respect to $\delta_M$ is
a \textsc{Knapsack} instance where the size of the knapsack is $c(v)$.
%
If weights are polynomially-bounded, then $\delta_M(e)$ is bounded for
every arc $e \in A$, and therefore this instance of \textsc{Knapsack}
can be solved in polynomial time using dynamic programming.

\begin{theorem}
Algorithm~\ref{alg:grd} is a $\half$-approximation algorithm for \gcp,
if edge weights are integral and polynomially bounded.
\end{theorem}

Using stadard scaling and rounding we obtain the following result.

\begin{theorem}
There exists a $(\half-\eps)$-approximation algorithm for \gcp,
for every $\eps \in (0,\half)$.
\end{theorem}

### 2017-02-10 src/group-sub.tex
merge dror

Consider the submodular maximization formulation given in Section~\ref{sec:sub}.
Figure~\ref{fig:not submodular} shows that the function $w : 2^V \to \R$ is not
submodular anymore.

\begin{figure}
\begin{center}
\input{fig-not-submodular}
\end{center}
\caption{Consider the following instance where node capacity are
written inside nodes, arcs are labeled with their size and all weights
are equal to 1.  It is easy to see that if $f(S)$ is the optimal
solution where $S$ is the set of drivers, then $f(A) + f(B) = 2 \leq 3
= f(A \cup B) + f(A \cap B)$, thus, $f$ is not submodular.}
\label{fig:not submodular}
\end{figure}

>>>>

Consider the submodular maximization formulation given in
Section~\ref{sec:sub}, namely $\barw: 2^V \to \R$, where
$\barw(S) \eqdf w(M(S))$ and $M(S)$ is the maximum weight carpool
matching that satisfies $D_{M(S)} \subseteq S \subseteq V \setminus
P_{M(S)}$.
%
Figure~\ref{fig:not submodular} contains an instance that shows that
the function $\barw$ is not submodular anymore.

\begin{figure}
\begin{center}
\input{fig-not-submodular}
\end{center}
\caption{Consider the following instance where node capacity are
written inside nodes, arcs are labeled with their size and all weights
are equal to 1.  We have that $f(A) + f(B) = 2 < 3 = f(A \cup B) +
f(A \cap B)$.}
\label{fig:not submodular}
\end{figure}

### 2017-02-10 src/group.tex
merge dror

We now consider a variant of the problem when, in addition to the
capacity function $c:V \to \N$, and the weight (profit) function
$w:A \to \R$, we also have a size function $s:V \to \N$, and the
constraint $\din[M] \leq c(v)$ is replaced with the constraint
$\sum_{(u,v)}s(u) \leq c(v)$.  The motivation for this variant of
the problem is when we want to allow a group of people to travel
together in the same car, thus $s$ is the size of the group.  We start
by showing that this variant of the problem does not fit the
submodular maximization formulation anymore. We show, however, that
the problem still has a $(2 +\eps)$-approximation algorithm.  Finally,
we argue that the local search algorithm can be trivially generalized
to handle the new variant of the problem while keeping the same
approximation guarantees.

>>>>

We now consider \gcp which a variant of \carpool in which in addition
to the capacity function $c:V \to \N$, and the weight function
$w:A \to \R^+$, we also have a size function $s:V \to \N$, and the
constraint $\din[M](v) \leq c(v)$ is replaced with the constraint
$\sum_{u:(u,v) \in M} s(u) \leq c(v)$.
%
%The motivation for this variant of the problem is when we want to
%allow a group of people to travel together in the same car, thus $s$
%is the size of the group.
%
We start by showing that this variant of the problem does not fit the
submodular maximization formulation anymore.  We show, however, that
the problem still has a $(\half -\eps)$-approximation algorithm by
extending the local improvement algorithm from
Section~\ref{sec:improve}.
%
Finally, we show that the local search algorithm from
Section~\ref{sec:local} can be generalized to handle \gcp while
keeping the same approximation guarantees.

### 2017-02-10 src/improve.tex
merge dror

In this subsection we give a local search $(2+\eps)$-approximation
algorithm for \carpool.

We need a few definitions before presenting our algorithm.
%
Let $M$ be a feasible carpool matching.  The weight $w_M(v)$ of a
vertex $v$ with respect to $M$ is the sum of the weights of the arcs
in $M$ that are incident on $v$, namely
\[
w_M(v) \eqdf \sum_{(u, v) \in M} w(u, v) + \sum_{(v, u) \in M} w(v, u)
~.
\]
For a subset of vertices $U \subseteq V$ we define
$w_M(U) \eqdf \sum_{v \in U} w_M(v)$.

@@

\begin{observation}
\label{lm:val-twice}
$\sum_{v \in V} w_M(v) = 2 \sum_{e \in M} w(e)$.
\end{observation}
\begin{proof}
\(
\sum_{v \in V} w_M(v)
= \sum_{v \in V} \sum_{(u, v) \in M} w(u, v) +
    \sum_{v \in V} \sum_{(v, u) \in M} w(v, u) 
= 2 \sum_{e \in M} w(e)
\).
\end{proof}

@@

Given an arc $(u,v)$, let $\inc(u,v)$ be the set of arcs that incident
$(u, v)$, namely
\[
\inc(u,v) = (N^-(u) \times \{u\}) \cup (\{v\} \times N^+(v))
~.
\]
If $S$ is a set of arcs, then $\inc(S) \eqdf \bigcup_{(u,v) \in
S} \inc(u,v)$.
%
Figure~\ref{fig:defs} depicts all the above definitions.

@@

We are now ready to describe our local search algorithm
(Algorithm~\ref{alg:grd}).  The algorithm starts with an empty
carpool matching $M$, and in every iteration it looks for a vertex
that can be improved.  If there exists such a vertex, then the
algorithm removes the arcs that are incident on it from $M$, and adds
the arcs that improve the vertex.  The algorithm terminates when no
vertex can be improved.  Figure~\ref{fig:improvement} depicts an
improvement step.

\begin{algorithm}
\caption{
\label{alg:grd}
Local Search
}
$M \gets \emptyset$ \\
\Repeat{$\text{done}$}{
	$\text{done} \leftarrow \textsc{True}$ \\
	\For{$v \in V$}{
		\If{$\exists$~improvment~$S_v$}{
			$M \leftarrow M \setminus \inc(S_v) \cup S_v$	\\
			$\text{done} \leftarrow \textsc{False}$ \\
		}
	}
}
\end{algorithm}

@@

For a vertex $v$ and a set $S$ of edges entering $v$, let $N^-_S(v)
= \set{u : (u,v) \in S}$ be the set in-neighbors corresponding to $S$.

\begin{lemma}
\label{lm:no improve}
Let $M$ be a matching computed by the local search algorithm.  Let $v$
be a vertex with no improvement, and let $S \subseteq N^-(v)$, such
that $\abs{S} \leq c(v)$, then $w(S) \leq w_M(v) + w_M(N^-_S(v))$.
\end{lemma}
\begin{proof}
If no improvement exists, then
\[
w(S) - w_M(N^-_S(v))
=    \sum_{(u,v) \in S} (w(u,v) - w_M(u))
=    \delta_M(S) 
\leq w_M(v)
~.
\]
\end{proof}

@@

\begin{lemma}
If Algorithm~\ref{alg:grd} terminates, then the computed solution is a
$2$-approximation.
\end{lemma}
\begin{proof}
Let $M$ be the matching produced by the algorithm, and let $M^*$ be an
optimal matching.  We load every vertex $v$ with an amount of money
equal to $w_M(v)$, and then we show that this is enough to pay for
every arc in the optimal matching.  Due to
Observation~\ref{lm:val-twice} the total amount of money that we use
is exactly twice the weight of $M$.

Consider a driver $v \in D_{M^*}$, and let $S = (V \times \{v\}) \cap
M^*$.  By lemma~\ref{lm:no improve} we know that $w(S) \leq w_M(v) +
w_M(N^-_S(v))$, thus we can pay for $S$, using the money on $v$ and on
$N^-_S(v)$.  Clearly, these vertices will not be charged again.
\end{proof}

@@

\begin{figure}
\centering
\input{fig-2-tight}
\caption[]{
\label{fig:grd worst}
GRD algorithm, worst case example: \\
Consider a path with $2n + 1$ arcs,
and alternating arc weights (2 and 1),
if the GRD algorithm selects all the 1 weighted arcs,
then no further improvement can be done and the value of the matching is $n + 1$,
while the optimal matching has value of $2n$.
}
\end{figure}

@@

\begin{theorem}
\label{thm:ls}
Algorithm~\ref{alg:grd} is a $2$-approximation algorithm for \carpool,
if edge weights are integral and polynomially bounded.
\end{theorem}
\begin{proof}
First, observe that determining if a vertex $v$ can be improved can be
done efficiently by considering the incoming arcs to $v$ in a
non-increasing order of their $\delta$s, and only ones with positive
values.  A vertex $v$ can be improved, then, if the $\delta$s of the
first $c(v)$ (or less) arcs sum up to more than $w_M(v)$.
%
It follows that the running time of an iteration of the for-loop is
polynomial.  Since the edge weights are integral and polynomially
bounded, the weight of an optimal carpool matching is polynomially
bounded.  The algorithm runs in polynomial time, because in each
iteration the algorithm improves the weight of the matching by at
least one or otherwise it terminates.
\end{proof}

It remains to consider the general case, where weight are not
polynomially bounded.  In this case one can use scaling and rounding
to ensure a polynomial running time in the cost of a $(1+\eps)$ factor
in the approximation ratio.

\begin{theorem}
There exists a $(2+\eps)$-approximation algorithm for \carpool, for
every $\eps > 0$.
\end{theorem}
\begin{proof}
Let $\eps' = \eps/4$.
%
Define the weight function $w'(e)
= \floor{\frac{w(e)}{w_{\max}} \cdot \frac{m}{\eps'}}$.
%
Let $M^*$ and $M'$ be an optimal carpool matchings with respect to $w$
and $w'$, resp.  We have that
\[
w'(M^*)
=    \sum_{e \in M^*} \floor{\frac{w(e)}{w_{\max}} \cdot \frac{m}{\eps'}}
>    \sum_{e \in M^*} \paren{ \frac{w(e)}{w_{\max}} \cdot \frac{m}{\eps'} } - m
=    \frac{m}{\eps' w_{\max}} \cdot w(M^*) - m
~.
\]
Since $w(M^*) \geq w_{\max}$, we have that 
\[
w'(M^*)
>    \frac{m}{\eps' w_{\max}} \cdot (w(M^*) - \eps' w_{\max})
\geq \frac{m}{\eps' w_{\max}} \cdot (1-\eps') w(M^*)
~.
\]
By Theorem~\ref{thm:ls}, our local search algorithm computes a
$2$-approximate carpool matching $M$ on $(G, c, w')$, and this
matching satisfies $w'(M) \geq w'(M')/2$.  Furthermore, since
$w'(M') \geq w'(M^*)$, it follows that $w'(M) \geq w'(M^*)/2$.
Therefore
\[
w(M)
\geq \frac{\eps' w_{\max}}{m} w'(M) 
\geq \frac{1}{2} \frac{\eps' w_{\max}}{m} w'(M^*)
>    \frac{1-\eps'}{2} w(M^*)
\geq \frac{w(M^*)}{2+\eps}
~.
\]
\end{proof}

>>>>

In this subsection we give a local improvement
$(\half-\eps)$-approximation algorithm for \carpool.  This algorithm
repeatedly checks whether the current carpool matching $M$ can be
improved by means of a star centered at a vertex $v$.  The profit from
this star is the total weight of the arcs in the star, and the cost is
the total weight of lost arcs (e.g., arcs from passengers to drivers
that became passengers of $v$).  If the profit is larger than the
cost, then an improvement step is performed.  The algorithm terminates
when such a step is not possible.
%
We remind the reader that this algorithm will be extended to \gcp in
Section~\ref{sec:group}.

We need a few definitions before presenting our algorithm.
%
Given a directed graphs $G = (V,A)$, define $\nin[] \eqdf \set{u :
(u,v) \in A}$ and $\nout[] \eqdf \set{u : (v,u) \in A}$.
%
Let $M$ be a feasible carpool matching.  The weight $w_M(v)$ of a
vertex $v$ with respect to $M$ is the sum of the weights of the arcs
in $M$ that are incident on $v$, namely
\[
\textstyle
w_M(v)
\eqdf w(M \cap \nin[]) + w(M \cap \nout[])
=     \sum_{(u, v) \in M} w(u, v) + \sum_{(v, u) \in M} w(v, u)
~.
\]
For a subset of vertices $U \subseteq V$ we define
$w_M(U) \eqdf \sum_{v \in U} w_M(v)$.

@@

\begin{observation}
\label{lm:val-twice}
$w_M(V) = 2 w(M)$.
%\sum_{v \in V} w_M(v) = 2 \sum_{e \in M} w(e)$.
\end{observation}
\begin{proof}
\(
\sum_{v \in V} w_M(v)
= \sum_{v \in V} \sum_{(u, v) \in M} w(u, v) +
    \sum_{v \in V} \sum_{(v, u) \in M} w(v, u) 
= 2 \sum_{e \in M} w(e)
\).
\end{proof}

@@

Given an arc $(u,v) \in A$, let $\inc(u,v)$ be the set of arcs that
incident $(u, v)$, namely define
\(
\inc(u,v) \eqdf (\nin[](u) \times \{u\}) \cup (\{v\} \times \nout[](v))
%~.
\).
If $S$ is a set of arcs, then $\inc(S) \eqdf \bigcup_{(u,v) \in
S} \inc(u,v)$.
%
Figure~\ref{fig:defs} depicts all the above definitions.

@@

We are now ready to describe our local improvement algorithm
(Algorithm~\ref{alg:grd}).  The algorithm starts with an empty carpool
matching $M$, and in every iteration it looks for a vertex that can be
improved.  If there exists such a vertex $v$, then the algorithm
removes the arcs that are incident on it from $M$, and adds the arcs
in $S_v$.  The algorithm terminates when no vertex can be improved.
Figure~\ref{fig:improvement} depicts an improvement step.

\begin{algorithm}
\caption{Local Improvement}
\label{alg:grd}
$M \gets \emptyset$ \\
\Repeat{$\text{done}$}{
	$\text{done} \leftarrow \textsc{True}$ \\
	\For{$v \in V$}{
		\If{there exists an improvement $S_v$}{
			$M \leftarrow M \setminus \inc(S_v) \cup S_v$	\\
			$\text{done} \leftarrow \textsc{False}$ \\
		}
	}
}
\end{algorithm}

@@

For a vertex $v$ and a set $S$ of edges entering $v$, let $\nin[S](v)
= \set{u : (u,v) \in S}$ be the set in-neighbors corresponding to $S$.

\begin{lemma}
\label{lm:no improve}
Let $M$ be a matching computed by the local improvement algorithm.
Let $v$ be a vertex with no improvement, and let
$S \subseteq \nin(v)$, such that $\abs{S} \leq c(v)$, then $w(S) \leq
w_M(v) + w_M(\nin[S](v))$.
\end{lemma}
\begin{proof}
If no improvement exists, then we have that \\
\(
w(S) - w_M(\nin[S](v))
=    \sum_{(u,v) \in S} (w(u,v) - w_M(u))
=    \delta_M(S) 
\leq w_M(v)
~.
\)
\end{proof}

@@

\begin{lemma}
If Algorithm~\ref{alg:grd} terminates, then the computed solution is a
$\half$-approximation.
\end{lemma}
\begin{proof}
Let $M$ be the matching produced by the algorithm, and let $M^*$ be an
optimal matching.  We load every vertex $v$ with an amount of money
equal to $w_M(v)$, and then we show that this is enough to pay for
every arc in the optimal matching.  Due to
Observation~\ref{lm:val-twice} the total amount of money that we use
is exactly twice the weight of $M$.

Consider a driver $v \in D_{M^*}$, and let $S = (V \times \{v\}) \cap
M^*$.  By lemma~\ref{lm:no improve} we know that $w(S) \leq w_M(v) +
w_M(\nin[S](v))$, thus we can pay for $S$, using the money on $v$ and
on $N^-_S(v)$.  Clearly, these vertices will not be charged again.
\end{proof}

@@

\begin{figure}
\centering
\input{fig-2-tight}
\caption[]{%
Consider a path with $2n + 1$ arcs, and alternating arc weights (2 and
1), if the local improvement algorithm selects all arcs of weight 1,
then no further improvement can be done and the value of the matching
is $n + 1$, while the optimal matching has value of $2n$.}
\label{fig:grd worst}
\end{figure}

@@

\begin{theorem}
\label{thm:improve-bounded}
Algorithm~\ref{alg:grd} is a $\half$-approximation algorithm
for \carpool, if edge weights are integral and polynomially bounded.
\end{theorem}
\begin{proof}
First, observe that determining if a vertex $v$ can be improved can be
done efficiently by considering the incoming arcs to $v$ in a
non-increasing order of their $\delta_M$s, and only ones with positive
values.  A vertex $v$ can be improved, then, if the $\delta$s of the
first $c(v)$ (or less) arcs sum up to more than $w_M(v)$.
%
It follows that the running time of an iteration of the for-loop is
polynomial.  Since the edge weights are integral and polynomially
bounded, the weight of an optimal carpool matching is polynomially
bounded.  The algorithm runs in polynomial time, because in each
iteration the algorithm improves the weight of the matching by at
least one or otherwise it terminates.
\end{proof}

It remains to consider the case, where weight are not polynomially
bounded.  In this case one can use scaling and rounding to ensure a
polynomial running time in the cost of a $(1+\eps)$ factor in the
approximation ratio.
%
The proof of the next theorem is given in the appendix.

\begin{theorem}
\label{thm:improve}
There exists a $(\half-\eps)$-approximation algorithm for \carpool,
for every $\eps \in (0,\half)$.
\end{theorem}

### 2017-02-10 src/intro.tex
merge dror

In this paper we assume that potential passenger satisfactions are
given as input and the goal is to compute an assignment of passengers
to driver so as to maximize the global satisfaction.
%
More formally, we are given a directed graph $G = (V, A)$, where each
vertex $v \in V$ corresponds to a user of the service, and an arc $(u,
v)$ exists if the user corresponding to vertex $u$ is willing to
commute with the user corresponding to vertex $v$.  We are given a
capacity function $c: V \rightarrow \N$ which bound the number of
passengers each user can drive if she was selected as a driver.  A
non-negative weight function $w : A \rightarrow \R^*+$ is used to model
the amount of satisfaction $w(u,v)$ that user $u$ gains when riding
with user $v$.
%
If $(u,v) \in A$ implies that $(v,u) \in A$ and $w(u,v) = w(v,u)$, we
call the instance \emph{undirected}.  If $w(v,u) = 1$, for every
$(v,u) \in A$, then we call the instance
\emph{unweighted}.  If $c(v) = \infty$, for every $v$, then we
call the instance \emph{uncapacitated}.

Given a directed graph $G$ and a subset $M \subseteq A$, define
$\din[M](v) \eqdf \abs{\set{u :(u,v) \in M}}$ and
$\dout[M](v) \eqdf \abs{\set{u :(v,u) \in M}}$.
%
A feasible \emph{carpool matching} is a subset of arcs, $M \subseteq
A$, such that every $v \in V$ satisfies:%
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A feasible carpool matching $M$ partition of $V$ as follows:
\begin{align*}
P_M & \eqdf \set{v : \dout[M](v) = 1} \\
D_M & \eqdf \set{v : \din[M](v) \geq 1} \\
Z_M & \eqdf \set{v : \dout[M](v) = \din[M](c) = 0}
~.
\end{align*}
where $P_M$ is the set of \emph{passengers}, $D_M$ is the set of
\emph{active drivers}, and $Z_M$ is the set of \emph{solo drivers}.
%
In the \carpool problem the goal is to find a matching $M$ of maximum
total weight, namely to maximize $w(M) \eqdf \sum_{(v,u) \in M}
w(v,u)$.  In other words, the \carpool problem is about finding a set
of (directed toward the center) vertex disjoint stars that maximizes
the total weight of the arcs.  
%
Figure~\ref{fig:carpool} is an example of the \carpool problem.
%
Note that in the unweighted case the goal is to find a carpool
matching $M$ that maximizes $\abs{P_M}$.
%
Moreover, observe that if $G$ is undirected, $D_M \cup Z_M$ is
a \emph{dominating set}.  Hence, in this case, an optimal carpool
matching induces an optimal dominating set and vice versa.
Since \textsc{Minimum Dominating Set} is NP-hard, if follows
that \carpool is NP-hard even if the instance is undirected,
unweigthed, and uncapacitated.

@@

\subparagraph{Related work.}
%
Agatz et al.~\cite{agatz2012optimization} outlined the optimization
challenges that arise when developing technology to support
ride-sharing and survey the related operations research models in the
academic literature.
%
Hartman et al.~\cite{hartman2014theory} designed
several heuristic algorithms for the \carpool problem and compared
their performance on real data.  Other heuristic algorithms were
developed as well~\cite{knapen2014exploiting}.
%
Hartman~\cite{hartman2013optimal} proved that the \carpool problem is
NP-hard even in the case where the weight function is binary and
$c(v) \leq 2$ for every $v \in V$.  In addition, Hartman presented a
natural integer linear programs and showed that if the set of drivers
is known, then an optimal assignment of passengers to drivers can be
found in polynomial time using a reduction to \textsc{Network Flow}
(see also Kutiel~\cite{kutiel2016}.)

@@

@@

Another related problem is \textsc{$k$-Set Packing}.  In this problem
one is given a collection of weighted sets, each containing as most
$k$ elements, and the goal is to find a maximum weight subcollection
of disjoint sets.  Chandra and Halld\'orsson~\cite{chandra2001greedy}
presented a $\frac{3}{2(k+1)}$-approximation algorithm for this
problem.

Consider a subset of
nodes $U$ of size at most $k$.  Observe that each subset of nodes has
an optimal internal assignment of passenger to drivers.  Let the
weight of this assignment be the profit of $U$, denote $p(U)$.  If $k
= O(1)$, then $p(U)$ can be computed for every $U$ of size at most $k$
in polynomial time.  The outcome is a \textsc{Set Packing} instance.

Arkin et al.~\cite{arkin2004approximations} considered other variants
of capacitated star packing where a capacity vector is given as part
of the input and capacities need to be assigned to vertices.

Kutiel~\cite{kutiel2016} presented a $\frac{1}{3}$-approximation
algorithm for the problem and $\frac{1}{2}$-approximation algorithm for the
unweighted variant of the problem.

@@

\subparagraph{Our contribution.}
%
In Section~\ref{sec:sub} we show that the \carpool can be formulated
as unconstrained submodular maximization problem, thus it has a
$\frac{1}{2}$-approximation algorithm due
to~\cite{BFNS15,buchbinder2016deterministic}.

In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide a $(\frac{\cmax+1}{2\cmax} - \eps)$-approximation
algorithm, for any $\eps>0$, where $\cmax \eqdf \max_{v \in V} c(v)$
is the maximum capacity.  We also show that \carpool is APX-hard even
if $\cmax \leq 3$.
%
In Section~\ref{sec:group} we consider \gcp.  
We first show that the unconstrained submodular maximization formulation for
the \carpool does not follow for the \gcp.
We show, however, that this problem still admits
a $(\frac{1}{2} -\varepsilon)$-approximation algorithm, and argue that the local
search algorithm generalizes to \gcp as well.

>>>>

%In this paper
We assume that potential passenger satisfactions are given as input
and the goal is to compute an assignment of passengers to driver so as
to maximize the global satisfaction.
%
More formally, we are given a directed graph $G = (V, A)$, where each
vertex $v \in V$ corresponds to a user of the service, and an arc $(u,
v)$ exists if the user corresponding to vertex $u$ is willing to
commute with the user corresponding to vertex $v$.  We are given a
capacity function $c: V \rightarrow \N$ which bound the number of
passengers each user can drive if she was selected as a driver.  A
non-negative weight function $w : A \rightarrow \R^+$ is used to model
the amount of satisfaction $w(u,v)$ that user $u$ gains when riding
with user $v$.
%
If $(u,v) \in A$ implies that $(v,u) \in A$ and $w(u,v) = w(v,u)$, we
call the instance \emph{undirected}.  If $w(v,u) = 1$, for every
$(v,u) \in A$, then we call the instance
\emph{unweighted}.  If $c(v) = \infty$, for every $v$, then we
call the instance \emph{uncapacitated}.

Given a directed graph $G$ and a subset $M \subseteq A$, define
$\din[M](v) \eqdf \abs{\set{u :(u,v) \in M}}$ and
$\dout[M](v) \eqdf \abs{\set{u :(v,u) \in M}}$.
%
A feasible \emph{carpool matching} is a subset of arcs, $M \subseteq
A$, such that every $v \in V$ satisfies:%
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A feasible carpool matching $M$ partitions of $V$ as follows:
\begin{align*}
P_M & \eqdf \set{v : \dout[M](v) = 1} \\
D_M & \eqdf \set{v : \din[M](v) \geq 1} \\
Z_M & \eqdf \set{v : \dout[M](v) = \din[M](c) = 0}
~.
\end{align*}
where $P_M$ is the set of \emph{passengers}, $D_M$ is the set of
\emph{active drivers}, and $Z_M$ is the set of \emph{solo drivers}.
%
In the \carpool problem the goal is to find a matching $M$ of maximum
total weight, namely to maximize $w(M) \eqdf \sum_{(v,u) \in M}
w(v,u)$.  In other words, the \carpool problem is about finding a set
of (directed toward the center) vertex disjoint stars that maximizes
the total weight of the arcs.  
%
Figure~\ref{fig:carpool} contains an example of a \carpool instance.
%
Note that in the unweighted case the goal is to find a carpool
matching $M$ that maximizes $\abs{P_M}$.
%
Moreover, observe that if $G$ is undirected, $D_M \cup Z_M$ is
a \emph{dominating set}.  Hence, in this case, an optimal carpool
matching induces an optimal dominating set and vice versa.
Since \textsc{Minimum Dominating Set} is NP-hard, if follows
that \carpool is NP-hard even if the instance is undirected,
unweigthed, and uncapacitated.

@@

\subparagraph{Related work.}
%
Agatz et al.~\cite{agatz2012optimization} outlined the optimization
challenges that arise when developing technology to support
ride-sharing and survey the related operations research models in the
academic literature.
%
Hartman et al.~\cite{hartman2014theory} designed
several heuristic algorithms for the \carpool problem and compared
their performance on real data.  Other heuristic algorithms were
developed as well~\cite{knapen2014exploiting}.
%
Hartman~\cite{hartman2013optimal} proved that the \carpool problem is
NP-hard even in the case where the weight function is binary and
$c(v) \leq 2$ for every $v \in V$.  In addition, Hartman presented a
natural integer linear programs and showed that if the set of drivers
is known, then an optimal assignment of passengers to drivers can be
found in polynomial time using a reduction to \textsc{Network Flow}
(see also Kutiel~\cite{kutiel2016}.)
%
Kutiel~\cite{kutiel2016} presented a $\frac{1}{3}$-approximation
algorithm for \carpool and a $\half$-approximation algorithm for the
unweighted variant of \carpool.

@@

Arkin et al.~\cite{arkin2004approximations} considered
the \textsc{Maximum Capacitated Star Packing} problem.  In this
problem the input consists of a complete undirected graph with
non-negative edge weights and a capacity vector $c
= \set{c_1,\ldots,c_p}$, where $\sum_{i=1}^p c_i = \abs{V} - p$.  The
goal is to find a set of vertex-disjoint stars in $G$ of size
$c_1,\ldots,c_p$ of maximum total weight.  Arkin et
al.~\cite{arkin2004approximations} provided a local search algorithm
whose approximation ratio is $\inv{3}$, and a matching-based
$\half$-approximation algorithm for the case where edge weights
satisfy the triangle inequality.

@@

Another related problem is \textsc{$k$-Set Packing}, where one is
given a collection of weighted sets, each containing as most $k$
elements, and the goal is to find a maximum weight subcollection of
disjoint sets.  Chandra and Halld\'orsson~\cite{chandra2001greedy}
presented a $\frac{3}{2(k+1)}$-approximation algorithm for this
problem.
%
\carpool can be seen a special case of \textsc{$k$-Set Packing}
with $k = \cmax + 1$.  Consider a subset of nodes $U$ of size at most
$k$.  Observe that each subset of nodes has an optimal internal
assignment of passenger to drivers.  Let the weight of this assignment
be the profit of $U$, denote $p(U)$.  If $k = O(1)$, $p(U)$ can be
computed for every $U$ of size at most $k$ in polynomial time.  The
outcome is a \textsc{$k$-Set Packing} instance.  This leads to a
$\frac{3}{2(\cmax+2)}$-approximation algorithm for the case where
$\cmax = O(1)$.

@@

\subparagraph{Our contribution.}
%
Sections~\ref{sec:approx} contains approximation algorithms
for \carpool.  First, in Section~\ref{sec:sub} we show that
the \carpool can be formulated as unconstrained submodular
maximization problem, thus it has a $\frac{1}{2}$-approximation
algorithm due to~\cite{BFNS15,buchbinder2016deterministic}.
%
We show a local improvement algorithm for \carpool%
\footnote{We use the term \emph{local improvement} to differentiate
this algorithm from the local search algorithm that is given in
Section~\ref{sec:local}.}.
%
This algorithm repeatedly checks whether the current carpool matching
can be improved by means of a star centered at a vertex, and it
terminates when such a step is not possible.
%
The approximation ratio of this algorithm is $\half$ if weights are
polynomially bounded, and its ratio is $\half-\eps$ in general.

Section~\ref{sec:cmax} we consider \carpool with bounded maximum
capacity.   Define $\cmax \eqdf \max_{v \in V} c(v)$.
%
In Section~\ref{sec:hardness} we show that \carpool is APX-hard even
for undirected and unweighted instances with $\cmax \leq b$, for any
$b \geq 3$.
%
In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide a local search algorithm whose approximation
ratio is $\half + \inv{2\cmax} - \eps$, for any $\eps>0$.

\dror{Compare to [7]!}

Finally, Section~\ref{sec:group} discusses \gcp.  We first show that
the unconstrained submodular maximization formulation for the \carpool
does not work for \gcp.  We show, however, that this problem still
admits a $(\frac{1}{2} -\varepsilon)$-approximation algorithm by
extending the local improvement algorithm.  In addition, we show that
the local search algorithm generalizes to \gcp with the same
approximation ratio.
